---
title: "PracMachineLearning"
author: "Geetika G"
date: "January 31, 2016"
output: html_document
---
# Background

This is the document for the practical machine learning assignment. In this assignment, I have loaded the data, created the partition sets, used the models and cross validated them to make sure the model with the best accuracy is chosen. I have then applied that model to the test data to apply those to the 20 scenarios and get the quiz responses.

Loading the required libraries and data

```{r}
library(tree)
library(caret)

setwd("/Users/geetikagangwani/Downloads/.")
getwd()
data_train <- read.csv("pml-training.csv", header = TRUE, na.strings = c("NA", "#DIV/0!", ""))
data_test <- read.csv("pml-testing.csv", header = TRUE, na.strings = c("NA", "#DIV/0!", ""))

dim(data_train)

```

Removing the information columns, not needed for our analysis

```{r}
data_train <- data_train[, 7:160]
data_test <- data_test[, 7:160]
```

Removing any data that has NA

```{r}
is_data  <- apply(!is.na(data_train), 2, sum) >19621
data_train <- data_train[, is_data]
data_test <- data_test[, is_data]
```

Partitioning the training data into training and testing. Using 70% of the data to be traingin and 30% to be testing.

```{r}
set.seed(10001)
inTrain = createDataPartition(y = data_train$classe, p = 0.7, list = FALSE)
sample_train <- data_train[inTrain, ]
sample_test <- data_train[-inTrain, ]
```

# Analysis

## 1. First using the rpart model to analyze the data

```{r}
modFit <- train(classe ~ ., method = "rpart", data = sample_train)
print(modFit$finalModel)
```

## Cross Validation
```{r}
pred_rpart = predict(modFit, sample_test)
confusionMatrix(pred_rpart,sample_test$classe)$overall["Accuracy"]
```

0.5 is not very accurate, hence we cannot rely on this model to calculate the result.

## 2. Random forest method


```{r}
require(randomForest)
set.seed(10002)
rf_train = randomForest(classe ~ ., data = data_train)
rf_train
plot(rf_train)
```


## Cross Validation
```{r}
rf_pred = predict(rf_train, sample_test)
confusionMatrix(rf_pred,sample_test$classe)$overall["Accuracy"]
```

This is a very accurate model and will be used to the testing data to predict the result

```{r}
quiz <- predict(rf_train, data_test)
quiz
```
